---
output: github_document
editor_options:
  markdown:
    wrap: 72
---

# scholid

[![R-CMD-check](https://github.com/Thomas-Rauter/scholid/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Thomas-Rauter/scholid/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://app.codecov.io/gh/Thomas-Rauter/scholid/branch/main/graph/badge.svg)](https://app.codecov.io/gh/Thomas-Rauter/scholid)

`scholid` provides lightweight, dependency-free utilities for working with
scholarly identifiers in R. The package is designed as a small, well-tested
foundation that can be safely reused by other packages and data workflows.

## Installation

Install the released version from CRAN:

```{r install, eval = FALSE}
install.packages("scholid")
```

## Scope

The package focuses on common identifier systems used in scholarly
communication:

- DOI
- ORCID iD
- ISBN
- ISSN
- arXiv
- PubMed (PMID)
- PubMed Central (PMCID)

## Interface

User-available functions:

| Function | Purpose |
|---------|---------|
| `scholid_types()` | List supported scholarly identifier types |
| `is_scholid(x, type)` | Test whether values conform to a given identifier type |
| `normalize_scholid(x, type)` | Normalize identifiers to canonical form |
| `extract_scholid(text, type)` | Extract identifiers of a given type from free text |
| `classify_scholid(x)` | Guess the identifier type of each input value |

## Examples

```{r example, eval = TRUE}
# list supported scholarly identifier types
scholid::scholid_types()

# test whether values match a given identifier type
scholid::is_scholid(
  x    = "10.1000/182",
  type = "doi"
)

# normalize identifiers to canonical form
scholid::normalize_scholid(
  x    = "https://doi.org/10.1000/182",
  type = "doi"
)

# extract identifiers of a given type from free text
scholid::extract_scholid(
  text = "See https://doi.org/10.1000/182 for details.",
  type = "doi"
)

# classify the identifier type of each input value
scholid::classify_scholid(
  x = c(
    "10.1000/182",
    "0000-0002-1825-0097",
    "not an id"
  )
)
```

For more detailed usage patterns, including extraction from text and
classification of mixed identifier columns, see the **Get started**
vignette.

## License

MIT
